\documentclass[11pt]{article}
\usepackage[top=1.00in, bottom=1.0in, left=1in, right=1in]{geometry}
\renewcommand{\baselinestretch}{1.1}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{parskip}
\usepackage{xcolor}
\usepackage[normalem]{ulem}

\def\labelitemi{--}
\parindent=0pt

\begin{document}
\renewcommand{\refname}{\CHead{}}


\title{Supplement: The problem and promise of \\ modeling plant leafout under warmer winters} %  How overconfident models of tree leafout have stalled progress... Infinite models of leafout given warming winters...? % I don't love this title.
\author{E.M. Wolkovich$^1$, Justin Ngo$^1$, Victor van der Meersch$^1$, Jonathan Auerbach$^2$} % also possibly: Rob Guy -- I emailed him in mid-December 2025, have not heard back
\date{\today}
\maketitle

$^1$ Forest \& Conservation Sciences, Faculty of Forestry, University of British Columbia, 2424 Main Mall, Vancouver, BC V6T 1Z4, Canada\\
$^2$ Department of Statistics, George Mason University, 4511 Patriot Circle, Fairfax, VA 22030, United States\\

\emph{The search for better indicators}

Increasing evidence suggests that callose and its degradation through 1,3-$\beta$-glucanases may control bud endodormancy \citep{rinne2011,vanderschoot2014,pan2023epigenetic,pandey2026variable}, potentially alongside phytohormones \citep[][]{tylewicz2018photoperiodic,pan2021aba}---but how useful any of these will be as measurable indicators is unclear. 1,3-$\beta$-glucanases localized to the plasmodesmata have displayed upregulated expression during the overwintering/endodormancy period. Thus, measuring callose deposition and degradation in the bud may be possible through this upregulation, though this is far from simple. Phytohormones, such as ABA and GA, have also been repeatedly implicated and have similarly measurable counterparts. ABA levels increase during dormancy induction and decrease during dormancy release, but do not yet provide a clear indicator. 

Finding a particular transcript or ratio of transcripts would be likely far easier to measure than callose-related compounds, ABA or gibberellin. One possibility may be in some of the dormancy -associated MADS-box genes (DAM), which are important central regulators of bud dormancy in perennial species \citep{falavigna2019want,yang2021bud}. \citet{zhang2023ppmapk6} have recently shown that during endodormancy in peaches, ABA promotes the expression of PpDAM6 while simultaneously inhibiting the expression of an interacting mitogen-activated protein kinase (PpMAPK6). The expression of PpMAPK6 increases as ABA content decreases during the chilling-induced transition from endodormancy to ecodormancy. Phosphorylation of PpDAM6 by PpMAPK6 may then accelerate its degradation and promote dormancy release. ABA also mediates callose deposition and plasmodesmata closure during dormancy development \citep{tylewicz2018photoperiodic}. After prolonged chilling, gibberellic acid (GA) biosynthesis is activated and the callose is degraded \citep{rinne2011,singh2018genetic}. 

% From Justin (11 Jan 2026): 
% Researcher could measure upregulation expression of 1,3-$\beta$-glucanases localized to the plasmodesmata during the overwintering/endodormancy period, providing a method of quantifying and visualizing callose deposition and degradation in the bud during chilling. Likewise, phytohormones heavily involved in dormancy induction and release like ABA and GA have similarly measurable counterparts, such as the ABA-mediated expressions of [MADS-box associated proteins/morphogenesis and development associated proteins/ PpDAM6] and [MAP kinases/protein activation and deactivation cascade-associated enzymes/PpMAPK6] in peaches.

% From Rob Guy (Jan 2026): 
% Yes, callose or the glucanases may be appropriate indicators. Ideally, though, it should be something easily measureable, like a particular transcript or ratio of transcripts (perhaps for the glucanase). Although ABA and gibberellin are implicated, they are not easy to measure. I'll attach some text I wrote for another purpose (but never used) if you want to flesh this out a bit.


\bibliographystyle{/Users/Lizzie/Documents/EndnoteRelated/Bibtex/styles/besjournals}
\section{References}
\bibliography{..//refs/bibforchillingconceptms}

\end{document}
